A generic and fast C++ optimization framework

### 2017-10-21 paper.tex
Fix example code issues, add output and SFINAE reference.

% \vspace*{-0.3em}
%\begin{listing}[H]
\begin{minted}[fontsize=\footnotesize]{c++}
class ObjectiveFunction {
  ObjectiveFunction() {
    in = arma::vec("20 12 15 100");
    bi = arma::vec("-4 -2 -3 -8");
  }

@@

  double Evaluate(const arma::mat& para, size_t s, size_t bs) {
    double cost = 0;
    for (size_t i = s; i < s + bs; i++)
      cost += (para(i) * para(i)) + bi(order(i)) * para(order(i)) + in(order(i));
    return cost;
  }

  void Gradient(const arma::mat& para, size_t s, size_t bs, arma::mat& g) {
    g.zeros(para.n_rows, para.n_cols);
    for (size_t i = s; i < s + bs; i++)
      g(i) += (1.0 / bs) * 2 * para(i) + bi(order(i));
  }

  arma::vec in, bi; arma::uvec order;
};
\end{minted}
%\vspace{-0.6em}
%\caption{Simple objective function where each dimension has a parabola with a
%  distinct minimum}
%\vspace{-0.8em}
%\end{listing}
%\vspace{-0.8em}
%\begin{listing}[H]
\begin{minted}[fontsize=\footnotesize]{c++}
class SimpleOptimizer {
  SimpleOptimizer(size_t bs = 2, double lr = 0.1) : bs(bs), lr(lr) { }

  template<typename FunctionType>
  double Optimize(FunctionType& function, arma::mat& parameter) {
    arma::mat gradient;
    for (size_t i = 0; i < 300; i += bs) {
      const size_t funIdx = i % 4;
      function.Gradient(parameter, funIdx, gradient, batchSize);
      parameter -= lr * gradient;
      if (funIdx == 0) { function.Shuffle(); }
    }
    return function.Evaluate(parameter, 0, function.NumFunctions());
  }
  size_t bs; double lr; // Store batch size and learning rate internally.
};
\end{minted}
%\vspace{-0.6em}
%\caption{Simple stochastic gradient descent method with a static number of
%iterations to optimize the function parameter ({\tt bs} defines the batch size
%and {\tt lr} the learning rate). )}
%\vspace{-0.6em}
%\end{listing}
%\vspace{-0.8em}
Using this code, we can compile the code, and when we run it

\vspace*{-0.4em}
\begin{minted}[fontsize=\footnotesize]{c++}
  ObjectiveFunction function; arma::mat parameter("0 0 0 0;")
  SimpleOptimizer optimizer;
  std::cout << optimizer.Optimize(function, parameter);
\end{minted}
\vspace*{-0.4em}

we receive the following output:

\vspace*{-0.4em}
{\footnotesize
\begin{verbatim}
xxxx
\end{verbatim}
}
\vspace*{-0.4em}

@@

Unfortunately, template metaprogramming can result in some very lengthy error
messages.  Therefore, we must be careful to ensure that a user is able to easily
debug a problem when they implement an objective function or gradient.  To
provide good error message output, we can use C++'s template metaprogramming
support to determine what methods a type has available.  Similarly, we can also
use static compile-time constants to denote the methods that are required by a
specific optimizer.  This is implemented via SFINAE \cite{sfinae}.  A {\tt
static\_assert()} is raised when a given objective function does not have the
required methods for a given optimizer.

>>>>

% \vspace*{-0.3em}
%\begin{listing}[H]
\begin{minted}[fontsize=\footnotesize]{c++}
struct ObjectiveFunction {
  ObjectiveFunction() {
    in = arma::vec("20 12 15 100");
    bi = arma::vec("-4 -2 -3 -8");
  }

@@

  double Evaluate(const arma::mat& para, size_t s, size_t bs) {
    double cost = 0;
    for (size_t i = s; i < s + bs; i++)
      cost += (para(i) * para(i)) + bi(order(i)) * para(i) + in(order(i));
    return cost;
  }

  void Gradient(const arma::mat& para, size_t s, arma::mat& g, size_t bs) {
    g.zeros(para.n_rows, para.n_cols);
    for (size_t i = s; i < s + bs; i++)
      g(i) += (1.0 / bs) * 2 * para(i) + bi(order(i));
  }

  arma::vec in, bi; arma::uvec order;
};
\end{minted}
%\vspace{-0.6em}
%\caption{Simple objective function where each dimension has a parabola with a
%  distinct minimum}
%\vspace{-0.8em}
%\end{listing}
%\vspace{-0.8em}
%\begin{listing}[H]
\begin{minted}[fontsize=\footnotesize]{c++}
struct SimpleOptimizer {
  SimpleOptimizer(size_t bs = 1, double lr = 0.02) : bs(bs), lr(lr) { }

  template<typename FunctionType>
  double Optimize(FunctionType& function, arma::mat& parameter) {
    arma::mat gradient;
    for (size_t i = 0; i < 5000; i += bs) {
      const size_t funIdx = i % 4;
      if (funIdx == 0) { function.Shuffle(); }
      function.Gradient(parameter, funIdx, gradient, bs);
      parameter -= lr * gradient;
    }
    return function.Evaluate(parameter, 0, function.NumFunctions());
  }
  size_t bs; double lr; // Store batch size and learning rate internally.
};
\end{minted}
%\vspace{-0.6em}
%\caption{Simple stochastic gradient descent method with a static number of
%iterations to optimize the function parameter ({\tt bs} defines the batch size
%and {\tt lr} the learning rate). )}
%\vspace{-0.6em}
%\end{listing}
%\vspace{-0.8em}
Using this code, we can compile the code, and when we run it

\vspace*{-0.4em}
\begin{minted}[fontsize=\footnotesize]{c++}
  ObjectiveFunction function; arma::mat parameter("0 0 0 0;");
  SimpleOptimizer optimizer;
  std::cout << "objective: " << optimizer.Optimize(function, parameter);
\end{minted}
\vspace*{-0.4em}

we receive the output: {\tt objective: 123.8 }

The final value of the objective function should be close to the optimal value,
which is the sum of values at the vertices of the parabolas.
\vspace*{-0.4em}

@@

Unfortunately, template metaprogramming can result in some very lengthy error
messages.  Therefore, we must be careful to ensure that a user is able to easily
debug a problem when they implement an objective function or gradient.  To
provide good error message output, we can use C++'s template metaprogramming
support to determine what methods a type has available.  Similarly, we can also
use static compile-time constants to denote the methods that are required by a
specific optimizer.  This is implemented via SFINAE
\cite{Vandevoorde2002sfinae}.  A {\tt static\_assert()} is raised when a given
objective function does not have the required methods for a given optimizer.
